---
title: "Assignment 3"
author: "Gracie White"
date: "5/19/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Whale Strikes and willingness to pay 
Data columns are: (1) “risk”: level of risk reduction, (2) “bid”: annual payment for the household, (3) “vote”: 1 is yes, 0 is no, (4)”NEP” : measure of environmental concern (feel free to consult the WhaleStrikes Group Project) 

```{r}
library(tidyverse)

whale_df <- read_csv("HW3Data.csv")

```

###1. Create a linear probability model that predicts a respondent’s probability of voting ‘yes’ on the ballot based on their age, income, NEP score, the risk reduction offered by the program, and the cost of the program to that respondent. Show the model and interpret the regression coefficients.

```{r}
whales_lm <- lm(vote ~ age + income + NEP + bid + risk, data = whale_df)
summary(whales_lm)

#saving the coefficient's in R's brain
int <- whales_lm$coefficients[1]
age_50 <- whales_lm$coefficients[2]
age_40 <- abs(whales_lm$coefficients[3])
age_60 <- abs(whales_lm$coefficients[4])
age_30 <- whales_lm$coefficients[5]
inc_one <- whales_lm$coefficients[6]
inc_poor <- whales_lm$coefficients[7]
inc_rich <- whales_lm$coefficients[8]
inc_veryrich <- whales_lm$coefficients[9]
nep <- whales_lm$coefficients[10]
bid <- abs(whales_lm$coefficients[11])
risk <- whales_lm$coefficients[12]




```

**Regression Model:**  
  
$Probability(Voting~Yes) = `r round(int, digits = 4)` + `r round(age_30, digits=4)`(Age~to~30) - `r round(age_40, digits=4)`(Age~to~40) + `r round(age_50, digits=4)`(Age~to~50) - `r round(age_60, digits=4)`(Age~to~60) + `r round(inc_one, digits=4)`(Income~One~Percent) + `r round(inc_poor, digits=4)`(Income~Poor) + `r round(inc_rich, digits=4)`(Income~Rich) + `r round(inc_veryrich,digits=4)`(Income~Very~Rich) + `r round(nep, digits=4)`(NEP) - `r round(bid, digits=4)`(Bid) + `r round(risk, digits=4)`(Risk~Reduction)$

**Coefficient Interpretation:**  

###2. Based on this regression, what can you say about the value of a single prevented whale death?
(Hint: Think about how risk reduction for all whales translates into the number of whale deaths
avoided)
Reducing the risk of whale strikes by 20% saves five whales every year. Assume ratio stays constant. So, reduction by 4% saves a single whale. We'll find the willingess to pay for vessel speed reduction programs of 0% and compare to the willingness to pay for vessel speed reduction of 4% to fing WTP for a single whale.     


Assume p(yes) is ($p = 0.7$), assume an age to 30, income rich, and the average NEP (38.366).
WTP:
$0.7 = `r round(int, digits = 4)` + `r round(age_30, digits=4)`(Age~to~30) + `r round(inc_rich, digits=4)`(Income~Rich) + `r round(nep, digits=4)`(NEP) - `r round(bid, digits=4)`(Bid) + `r round(risk, digits=4)`(0)$

```{r, echo=FALSE}
#average of the yes votes:
avg_vote <- mean(whale_df$vote)
#average of the NEP:
nep_avg <- mean(whale_df$NEP)

#bid if risk reduction is zero:
bid_zero <- (int + age_30 + inc_rich + (nep*nep_avg) + (risk*0) - avg_vote)/bid
#39.5

#Bid if risk is .04
bid_four <- (int + age_30 + inc_rich + (nep*nep_avg) + (risk*4) - avg_vote)/bid
#42.3

```


Value of a single whale = 42.3 - 39.5 = $2.8/SB resident

###3. Pick three arbitrary respondents. For each of those respondents, estimate their WTP for a VSR program offering 60% risk reduction. 

```{r, include=FALSE}

#Picking 3:
# Outcome - 22, 233, 447
# Create new dataframe with the random respondents information
vsr60 <- whale_df %>% 
  slice(22,233,447)
 
#Assume p(yes) = .7


```

Individual 22:  
$`r avg_vote` = `r round(int, digits = 4)` + `r round(age_40, digits=4)`(Age~to~40) + `r round(inc_poor, digits=4)`(Income~Poor) + `r round(nep, digits=4)`(NEP) - `r round(bid, digits=4)`(Bid) + `r round(risk, digits=4)`(60)$

Individual 233:  
$`r avg_vote` = `r round(int, digits = 4)` - `r round(age_60, digits=4)`(Age~to~60) + `r round(inc_rich, digits=4)`(Income~Rich) + `r round(nep, digits=4)`(NEP) - `r round(bid, digits=4)`(Bid) + `r round(risk, digits=4)`(60)$  

Individual 447:  
$`r avg_vote` = `r round(int, digits = 4)`- `r round(age_60, digits=4)`(Age~over~60) + `r round(nep, digits=4)`(NEP) - `r round(bid, digits=4)`(Bid)+ `r round(risk, digits=4)`(60)$  


```{r}


bid_22 <- (int+age_40+inc_rich+(nep*15)+(risk*60)-avg_vote)/bid
bid_233 <- (int+age_60+inc_poor+(nep*51)+(risk*60)-avg_vote)/bid
bid_447 <- (int+age_60+(nep*54)+(risk*60)-avg_vote)/bid

```

person 22 = -$266 <- whale hater
person 233 = $260
person 447 = $302

###4. Now repeat this process and estimate the mean WTP for a VSR program offering 60% risk reduction among Santa Barbara County households. 
Assume the probability of voting yes the average of voting yes ($p = 0.7$). Use the average income bracket (middle), the average age bracket (to 50), and average NEP (38.366) to find WTP: 

$`r avg_vote` = `r round(int, digits = 4)` + `r round(age_50, digits=4)`(Age~to~50) + `r round(nep, digits=4)`(NEP) - `r round(bid, digits=4)`(Bid) + `r round(risk, digits=4)`(60)$  


```{r}
# average bid
bid_sb <- (int + age_50 + (nep*nep_avg) + (risk*60) - avg_vote)/bid
#64.5



```


WTP = $64.5 among SB Households

###5. If there are 150,000 households in Santa Barbara County, estimate the total benefit of a VSR program to these households

Risk = 60%
cost = $7 million
Total Benefits:
$64.5*150,000 = 9,675,000 $


Net benefits (surplus):
$9,675,000 - 7,000,000 = 2,675,000$

###6. Yes, the benefits outweight the costs by over $2,000,000

###7. Suppose the price of a carbon credit is $Z per ton, at what price Z* will the shipping industry decide to adopt the VSR for purely self-interested reasons? (Note: if ships voluntarily reduce speed, then the $7 million cost (noted above) of implementing the VSR would be avoided)

Assume that for any ship transiting the Santa Barbara Channel, a speed reduction (that results in a 60% risk reduction) will cost the shipper $1,000, but will result in 20 fewer tons of CO2 emitted per transit.

20*Z = $1,000
Z = $50/ton

###8. Now suppose the carbon credit price is exactly Z*, so all ships voluntarily reduce speed to achieve the 60% risk reduction. Approximately how many whales would be saved by this? What would be the social value of allowing ships to enter the carbon trading market?


a) 15 whales would be saved by this program, assuming the risk reduction ratio to whales saved stays constant 
b) the bid of saving 15 whales without any costs, 60% risk VSR = $9,675,000
